Untwisting the Tetrahelix (v0.12)

### 2017-03-24 doc/TetrahelixUntwist.tex
latest changes

\begin{figure}[H]
  \label{fig:closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up (partly along axis)}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/VerticalCloseUp.png}
     \caption{BC Helix Close-up (orthogonal)}
\end{figure}

@@

In Figure \ref{fig:spudboy}, the closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in Section \ref{sec:equitetrabeam}.

@@

\begin{figure}[H]
  \label{railanglefig}
     \centering
     \includegraphics[width=0.7\textwidth]{figures/RailAngleGeometry.png}
     \caption{Rail Angle Geometry}
\end{figure}

@@

The $H_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho{bc}$ rather than $\theta$.
Intuitively we see an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c2\pi/3 $.
\begin{align*}
 c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\}...} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{algebra...} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}

@@

From this formulation it is easy to see that moving one vertex on a rail moves us $\rho_{bc}$ radians around a circle. Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
we can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices. We further demand that all edge lengths
be finite, as we are only interested in physically constructable tetrahelices.
By isomporphic we mean there is a one-to-one mapping between both
vertices and edges in the two tetrahelices.
One could consider various definitions of optimality for a
tetrahelix, but the most useful to us as robotocists extending the Tetrobot concept is to minimize the
maximum difference between any two edges.

@@

\begin{figure}[H]
  \label{untouchingrailfig}
     \centering
     \includegraphics[width=0.7\textwidth]{figures/UntouchingRail.png}
     \caption{Untouching Rail}
\end{figure}

@@

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius is negative up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

@@

\begin{figure}[H]
  \label{fig:unfolded}
     \centering
     \includegraphics[width=0.6\textwidth]{figures/EquitetrabeamUnfolded.JPG}
     \caption{Projection Diagram}
\end{figure}

In Figure \ref{fig:spudboy}, the furthest tetrahelix is the optimal equitetrabeam.
Figure \ref{fig:unfolded} shows a representation of an unfolded equitetrabeam.
Note that folding the outer edges of the paper upward or downward effectively
chooses the chirality of the equitetrabeam.

@@

\begin{figure}[H]
  \label{fig:spudboy}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices}
\end{figure}

@@

Starting twenty years ago, Sanderson\cite{sanderson1996modular},
Hamlin,\cite{TetrobotBook}, and others including
Lee\cite{lee2002dynamic} created a style of robotics based on changing
the lengths of members joined at the center of a joint, thereby
creating a connection to pure geometry. More recently NASA has
experimented with tensegrities\cite{NTRT}, a different point in the
same design spectrum. These fields create a need to explore the notion
of geometries changing over time, not generally considered directly by
pure geometry.

@@

\begin{theorem}
  By changing only the length of the members that change rails and make two nodes, you can dynamically untwist a tetrobot
  forming the Boerdijk-Coxeter configuration to the equitetrabeam which rests flat on the plane.
\end{theorem}

\begin{proof}
  Proof by our computer program that does this using Equation \eqref{eq:general}.
  See video on website\cite{readtetrahelix}.

>>>>

\begin{figure}[H]
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up (partly along axis)}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/VerticalCloseUp.png}
     \caption{BC Helix Close-up (orthogonal)}
  \label{fig:closeup}
\end{figure}

@@

\begin{figure}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices}
  \label{fig:series}
\end{figure}

Figure \ref{fig:series} displays a continuum of tetrahelices optimal in a certain sense, which is the result of this
paper. The closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in Section \ref{sec:equitetrabeam}.

@@

\begin{figure}[H]
     \centering
     \includegraphics[width=0.7\textwidth]{figures/RailAngleGeometry.png}
     \caption{Rail Angle Geometry}
  \label{railanglefig}
\end{figure}

@@

The $H_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho{bc}$ rather than $\theta$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c2\pi/3 $.
\begin{align*}
 c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\}...} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{algebra...} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}

@@

From this formulation it is easy to see that moving one vertex on a rail
($H_{BCcolored}(n,c)$ to $H_{BCcolored}(n+1,c)$ for any $n$ and $c$)
moves us $\rho_{bc}$ radians around a circle. Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
we can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices. We further demand that all edge lengths
be finite, as we are only interested in physically constructable tetrahelices.
By isomporphic we mean there is a one-to-one mapping between both
vertices and edges in the two tetrahelices.
One could consider various definitions of optimality for a
tetrahelix, but the most useful to us as robotocists extending the Tetrobot concept is to minimize the
maximum difference between any two edges, because the Tetrobot uses linear acutators of limit range of extension.

@@

\begin{figure}[H]
     \centering
     \includegraphics[width=0.7\textwidth]{figures/UntouchingRail.png}
     \caption{Untouching Rail}
  \label{untouchingrailfig}
\end{figure}

@@

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius $r_{\rho}$ is negative up until $\rho_{bc}$ where it is $0$, 
we opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

@@

\begin{figure}[H]
     \centering
     \includegraphics[width=0.6\textwidth]{figures/EquitetrabeamUnfolded.JPG}
     \caption{Projection Diagram}
  \label{fig:unfolded}
\end{figure}

In Figure \ref{fig:series}, the furthest tetrahelix is the optimal equitetrabeam.
Figure \ref{fig:unfolded} shows a representation of an unfolded equitetrabeam.
Note that folding the outer edges of the paper upward or downward effectively
chooses the chirality of the equitetrabeam.

@@

@@

Starting twenty years ago, Sanderson\cite{sanderson1996modular},
Hamlin,cite{TetrobotBook}, Lee\cite{lee2002dynamic}, and others
 created a style of robotics based on changing
the lengths of members joined at the center of a joint, thereby
creating a connection to pure geometry. More recently NASA has
experimented with tensegrities\cite{NTRT}, a different point in the
same design spectrum. These fields create a need to explore the notion
of geometries changing over time, not generally considered directly by
pure geometry.

@@

\begin{theorem}
  By changing only the length of the longer members that connect two distinct rails (the two-hop members), you can dynamically untwist a tetrobot
  forming the Boerdijk-Coxeter configuration to the equitetrabeam which rests flat on the plane.
\end{theorem}

\begin{proof}
  Proof by our computer program that does this using Equation \eqref{eq:general} applied to the
  7-tet Tetrobot/Glussbot.
